---
title: "Naive Bayes Breast Cancer"
format: html
editor: visual
---

## Dataset & Paper:

The ***Breast Cancer Wisconsin (Diagnostic)*** dataset is a renowned collection of data used extensively in machine learning and medical research. Originating from digitized images of fine needle aspirates (FNA) of breast masses, this dataset facilitates the analysis of cell nuclei characteristics to aid in the diagnosis of breast cancer. In this article, we delve into the attributes, statistics, and significance of this dataset.

[*A new classifier for breast cancer detection based on Naïve Bayesian*](https://www.sciencedirect.com/science/article/abs/pii/S0263224115002419)

### Libraries

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(e1071)
library(here)
library(ggcorrplot)
library(ggpubr)
library(dplyr)
library(janitor)
```

### Read in Data

```{r}
breast_cancer_data <- read_csv(here("data/wisconsin_breast_cancer_data.csv"))
```

### Correlation Matrix 

Let's test the assumption that there are features that are independent from each other.

```{r}
breast_cancer_data_cor <- breast_cancer_data %>%
  select(-id, -diagnosis, -...33) #dropping these 3 columns 
```

### Create a matrix from the data frame 

```{r}
breast_cancer_matrix <- as.matrix(breast_cancer_data_cor)
```

### Calculate the correlations using cor()

```{r}
breast_cancer_correlations <- cor(breast_cancer_data_cor)
```

### Plot the correlations 

```{r}
ggcorrplot(breast_cancer_correlations, type = "lower", lab = TRUE, lab_size = 1)
```

Visualizing correlations between 2 features

```{r}
ggplot(data = breast_cancer_data, aes(x = radius_mean, y = perimeter_mean)) +
  geom_point() +
  geom_smooth() 
```

### Create a Naive Bayes Model

```{r}
breast_cancer_mod <- naiveBayes(diagnosis ~., data = breast_cancer_data)
```

Predict Diagnosis in our breast cancer data set

```{r}
breast_cancer_data <- breast_cancer_data %>% 
  mutate(predicted_diagnosis = predict(breast_cancer_mod, newdata = .))
```

Create the confusion matrix

```{r}
breast_cancer_data %>%
  tabyl(diagnosis, predicted_diagnosis) %>%
  adorn_percentages("row") %>% 
  adorn_pct_formatting(digits = 2) %>%
  adorn_ns
```
